\textbf{International Transport costs:\\New Findings from modeling additive costs}\\Online Appendix (Not for Publication)

### 2021-09-14 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
Tavail agregation
Signed-off-by: Guillaume Daudin <gdaudin@mac.com>
actulisation chiffres modele

\textbf{ATTENTION AJUSTER LES FIGURES ET CHIFFRES A NOUVELLE CALIBRATION}

@@

\begin{table}[htbp]
  \centering
  \caption{Calibration (1)} \label{tab:calib_horsTC}
\begin{center}
	\input{donnees_calibration.tex}
\end{center}
{\parbox[l]{10cm}{ \vspace{4pt}\footnotesize{Notes: $f_x$ can take two values, depending on the share of exporters considered that might differ between the initial steady state (21\%, based on early 1980s' value) and the final steady state (39\%, based on 2000's value.), according to the references provided above. BEJK (2003) refers to \cite{BEJK-AER-03}.}}}
\end{table}

@@

The results are reported in Table \ref{tab:resultats_modele}, panel (1), by transport mode (for Air (Columns (A)-(B)) and Vessel (Columns (C)-(D) respectively. Specifically, we report the change in total transport costs (expressed in percentage points, and based on the raw numbers reported in Table \ref{tab:calib_TC}), decomposed in its two dimensions (additive and ad-valorem); as well as the welfare change, both in absolute and in relative terms. We also run a robustness check of our results by maintaining a fixed export cost value. In this case, the welfare changes are solely attributable to the reduction in the variable transport costs. These results are reported in Table \ref{tab:resultats_modele}, panel (2).

\begin{table}[htbp]
  \centering
  \caption{Trade liberalization: The role of additive costs} \label{tab:resultats_modele}
\begin{center}
\begin{tabular}{l||c|c||c|c}
\hline
Mode & \multicolumn{2}{|c||}{Air} & \multicolumn{2}{c}{Vessel} \\ \hline
& (A) & (B) & (C) & (D)  \\
Reduction in: & $\tau$ only & $\tau$ and $t$ &$\tau$ only & $\tau$ and $t$\\  \hline
%$TC_{1974}$ & 0.062 & 0.062 & 0.062 & 0.062 & 0.105 & 0.105 & 0.105 & 0.105\\
%$TC_{2019}$ &  0.028 & 0.028 & 0.028 & 0.028&  0.041 & 0.041 & 0.041 & 0.041\\
$\Delta TC$ (in pp) &-3.4 &-3.4 &-6.4 &-6.4   \\
$\Delta \tau$ (in pp) & -3.4	&-1.4&-6.4&	-3.1	 \\
$\Delta t/\widetilde{p}_x^{\text{fas}}$ (in pp) & 0&	-2.0&	0&	-3.3 \\ \hline \hline
\multicolumn{4}{l}{(1) Lower fixed export cost} \\ \hline
%$\Delta \mathcal{W}$ & 0.0202 &  0.0386 & 0.0362 & 0.0540 & 0.0354 & 0.0535 & 0.0467 & 	0.064 \\
$\Delta \mathcal{W}/\mathcal{W}$ (in \%) & 	\textbf{2.21}	&\textbf{3.35}	&	\textbf{2.87}&	\textbf{4.00 }  \\\hline
Variety effect (\%)&0.75 &-1.04	&-0.13	&-1.90 \\
Price effect (\%)&	1.46 &	4.44&		 3.00&	6.06 \\ \hline \hline
\multicolumn{4}{l}{(2) Constant fixed export cost} \\ \hline
$\Delta \mathcal{W}/\mathcal{W}$ (in \%) & 1.23&	2.39 & 2.18	&3.38 	 \\ \hline
Variety effect (\%)& -0.83 	&-2.60	&-1.32	& -3.12 \\
Price effect (\%)&2.08 &5.20	&3.57	& 6.83  \\
\hline
\end{tabular}
\end{center}
{\parbox[l]{10cm}{ \vspace{4pt}\footnotesize{Notes: TC = Transport Costs, expressed in percentage of the average export fas price. The reduction in transport costs is expressed in percentage points. In the world including additive costs, we simulate the observed reduction in $\tau$ and iterate on the reduced value of $t$ such that the simulated value of $t/\widetilde{p}_x^{\text{fas}}$ matches the empirical target.  }}}
\end{table}

In line with the literature, we find that trade liberalization induces welfare gains whatever the transport mode (and the associated magnitude of the transport costs decrease). The originality of our results rather dwells on comparing the welfare gains depending on the type of cost reduction. On that side, the welfare gains are higher when the trade cost reduction occurs through a decrease in the additive component. This conclusion stands in accordance with the analytical results of \cite{sorensen2014}. This is consistent with the analytical results of \cite{sorensen2014}, in a different exercise of trade liberalization though. \footnote{\cite{sorensen2014} compares the welfare effect of a trade costs reduction depending whether it is exclusively ad-valorem or exclusively additive, the comparison between both trade cost decreases being a criterion of equal impact on trade openness. Our quantitative model offers a more general and flexible set-up, as we can virtually consider all cases of trade liberalization. Based on our empirical estimates, we compare the effects of the same decrease in overall transport costs depending on 1) whether additive costs are present or not, and 2) when they are present, assuming that the trade cost decrease is achieved by a reduction of both additive and ad-valorem components.} Quantitatively, the welfare gains are sizeable. The reduction variable transport costs generates a relative welfare gain higher by around 1.25 percentage point when such a reduction occurs through the additive component (which accounts for 50 to 60\% of the total cost decrease, depending on the transport mode), rather than when it realizes through a decrease in iceberg costs only (Table \ref{tab:resultats_modele}, panel (1), Column (B) vs (A) for Air, (D) vs (C) for Vessel). \cite{Irrazabal_2015} also quantify the welfare gains of a trade cost decrease in presence of additive costs. In contrast to \cite{Irrazabal_2015} though, our empirical analysis allows us to simulate the effects of the reduction in transport costs actually observed in the data over the period 1974-2019, in all their dimensions (additive and ad-valorem). \smallskip

@@

To understand the changes induced by the additive costs, we start considering the benchmark model with only ad-valorem costs (Table \ref{tab:resultats_modele}, Columns (A) and (C)). As reported in Table \ref{tab:resultats_modele}, the reduction in the iceberg cost induces a negative variety effect, which is yet more than counteracted by a positive competitive effect. The reduction in the number of available varieties for the household stands in line with the insights of the canonical \cite{melitz} model. With the reduction in $\tau$, more firms can now export, but this also increases the expected average profit. This raises the productivity threshold to enter the domestic market. Accordingly, the number of domestic active firms reduces, even if a larger number of them is able to export. As a whole, the total number of varieties accessible to the consumer reduces, which reduces welfare everything else equal.

This detrimental effect is counteracted by a positive price competitiveness effect. Due to the tougher selection on the domestic market, the less productive firms no longer enter. As a result, the average productivity on the domestic market increases, which pushes the average productivity upwards. As reported in Table \ref{tab:resultats_modele}, the strength of the price competitiveness effect dominates, inducing positive welfare gains with the reduction in iceberg costs, consistently with the analytical results of \cite{melitz}.

What does it change when the reduction in international transport costs also occurs through a reduction in its additive component? As underlined by \cite{sorensen2014}, a reduction in the additive cost reduces the export-market price of the high-productivity exporters relative to the low-productive ones (as can be inferred from Equation (\ref{eq:px})). Within the exporters category, this favors the more productive firms, whose market share, e.g. revenues, increases relatively more (see Equation (\ref{eq:rx})). Everything else equal, this pushes the average productivity on the export market upwards. Unlike the ad-valorem costs, the reduction in additive costs induces a \textit{within-exporters selection effect}, in favor of the more productive firms. At the general equilibrium, this within-exporters selection effect also affects the \textit{classic} selection effect at the extensive margin at the heart of the \cite{melitz} model. As the expected export profit increases, it is necessary to be more productive to satisfy the free-entry condition. Everything else equal, the threshold productivity on the domestic market rises, as well as the average productivity. Accordingly, the competitiveness effect increases by more when the trade liberalization involves a reduction in the additive component.

The more negative variety effect also results from the above mechanism. Since the reduction in additive costs favors the more productive firms in terms of market share, entry on the export market is tougher leading to a much more modest increase in the share of exporters. As the selection on the domestic market is also more stringent, the mass of total active firms reduces by a larger amount. From the consumer side, this induces a more negative variety effect that is welfare-detrimental everything else equal. As for the pure-iceberg case though, the strength of the (larger) competitiveness effect still dominates, inducing positive welfare gains, and even larger when the trade liberalization involves a reduction in additive costs.

As mentioned above, our results dwell on the joined reduction in both the variable and the fixed export costs, based on the empirical evidence provided in Section  \ref{sec:basic_results} and in \cite{Lincoln_McCallum2018}. One might yet wonder about the welfare gains induced by the reduction in variable costs only, i.e. \textit{ceteris paribus} for a given value of the fixed export cost. The answer can be found in Table \ref{tab:resultats_modele}, Panel (2). In this case, the fixed export cost is kept constant at its initial value, the only structural change being the decrease in variable transport costs (either ad-valorem only or both additive and ad-valorem). Two results emerge. First, the welfare gains induced by the trade liberalization are lower in magnitude. While this is not a surprising result \textit{per se} (since we only consider the reduction in one source of international frictions, rather than two), this exercise yet delivers a quantification of the welfare gains associated with a reduction in the fixed export cost: They are substantial, by an amount between 0.6 to 1 percentage point depending on the transport mode. Second, the main message of the paper remains valid. The welfare gains of the trade liberalization are still higher when part of this structural change is achieved through a reduction in additive transport costs, by the same order of magnitude (+ 1.25 percentage point). Put it differently, the reduction in the fixed export cost does not seem to depend on the composition of the variable costs. In the end, the welfare gains are the highest when trade liberalization induces a combined reduction in both variable and fixed export costs, and when the reduction in the variable part implies a reduction in the additive component, as observed empirically.\smallskip

In line with \cite{sorensen2014} and \cite{Irrazabal_2015}, our overall theoretical results point out the importance of taking into the additive components of international trade costs. Additive costs have a substantial impact on the extensive margin of trade, by affecting the composition of the basket of exported goods. This, in turn, has important consequences in normative terms. Relying on our empirical estimates of transport costs over 1974-2019, we obtain that the 3.4\% reduction of variable costs in Air transport, combined to the reduction in the fixed export costs, has induced an 3.5\% increase in welfare, much higher than if only iceberg costs were modeled (2.4\%). A similar result applies with the observed 6.4\% in total variable costs in maritime transport along with the reduction in the fixed export costs, with a relative increase in welfare by 4\% (as compared to 3.38\% in a pure ad-valorem world). In a nutshell, we show that neglecting additive costs leads to a substantial underestimation of the welfare gains of trade liberalisation. While this result has been obtained in the canonical \cite{melitz} model, they call for integrating additive costs in New Trade models in view of better assessing the (disputed) question of the welfare gains of trade liberalization.

@@

From this, we deduce the recursive link: $\widehat{t}_{ist} = \widehat{t}_{is74} \times \exp(\gamma_t)$.
Given the constraint $t>0$, we then obtain the percentage change from 1974 from $\Gamma^{add}_{ist} = 100\frac{\widehat{t}_{ist}}{\widehat{t}_{ik74}} = 100\exp(\gamma_t)$. Again, it is independent of the product-origin country pair. We can thus rewrite the time-trend series for the additive transport cost component as:
\begin{equation}
\Gamma^{add}_t  = 100\exp(\gamma^{add}_t) \label{eq:tcadd_compoeffect}
\end{equation}

>>>>

@@

\begin{table}[htbp]
  \centering
  \caption{Calibration (1)} \label{tab:calib_horsTC}
\begin{center}
	\input{donnees_calibration.tex}
\end{center}
{\parbox[l]{13cm}{ \vspace{4pt}\footnotesize{Notes: $f_x$ can take two values, depending on the share of exporters considered that might differ between the initial steady state (21\%, based on early 1980s' value) and the final steady state (39\%, based on 2000's value.), according to the references provided above. BEJK (2003) refers to \cite{BEJK-AER-03}.}}}
\end{table}

@@

The results are reported in Table \ref{tab:resultats_modele}, panel (1), by transport mode (for Air (Columns (A)-(B)) and Vessel (Columns (C)-(D) respectively.\footnote{For sake of space saving, we only report the welfare consequences of trade liberalization, depending on the structure of transport costs, even if we refer to the positive effects (in terms of number of firms, threshold and average productivities, etc.) to explain the underlying mechanisms. The whole set of results is available to the authors upon request.} Specifically, we report the change in total transport costs (expressed in percentage points, and based on the raw numbers reported in Table \ref{tab:calib_TC}), decomposed in its two dimensions (additive and ad-valorem); as well as the welfare change, both in absolute and in relative terms. We also run a robustness check of our results by maintaining a fixed export cost value. In this case, the welfare changes are solely attributable to the reduction in the variable transport costs. These results are reported in Table \ref{tab:resultats_modele}, panel (2).

\begin{table}[htbp]
  \centering
  \caption{Trade liberalization: The role of additive costs} \label{tab:resultats_modele}
\begin{center}
\begin{tabular}{l||c|c||c|c}
\hline
Mode & \multicolumn{2}{|c||}{Air} & \multicolumn{2}{c}{Vessel} \\ \hline
& (A) & (B) & (C) & (D)  \\
Reduction in: & $\tau$ only & $\tau$ and $t$ &$\tau$ only & $\tau$ and $t$\\  \hline
%$TC_{1974}$ & 0.062 & 0.062 & 0.062 & 0.062 & 0.105 & 0.105 & 0.105 & 0.105\\
%$TC_{2019}$ &  0.028 & 0.028 & 0.028 & 0.028&  0.041 & 0.041 & 0.041 & 0.041\\
$\Delta TC$ (in pp) &-3.4 &-3.4 &-6.4 &-6.4   \\
$\Delta \tau$ (in pp) & -3.4	&-1.4&-6.4&	-3.1	 \\
$\Delta t/\widetilde{p}_x^{\text{fas}}$ (in pp) & 0&	-2.0&	0&	-3.3 \\ \hline \hline
\multicolumn{4}{l}{(1) Lower fixed export cost} \\ \hline
$\Delta \mathcal{W}/\mathcal{W}$ (in \%) & 	\textbf{2.50}	&\textbf{3.16}	&	\textbf{3.63}&	\textbf{4.26 }  \\\hline
Variety effect (\%)&1.36 &0.48	&-0.43	&-1.28 \\
Price effect (\%)&	1.14 &	2.66&		 4.08&	5.63 \\ \hline \hline
\multicolumn{4}{l}{(2) Constant fixed export cost} \\ \hline
$\Delta \mathcal{W}/\mathcal{W}$ (in \%) & 1.23&	2.18 & 2.39	&3.38 	 \\ \hline
Variety effect (\%)& -0.83 	&-1.32	&-2.60	& -3.12 \\
Price effect (\%)&2.08 &3.56	&5.20	& 6.83  \\
\hline
\end{tabular}
\end{center}
{\parbox[l]{10cm}{ \vspace{4pt}\footnotesize{Notes: TC = Transport Costs, expressed in percentage of the average export fas price. The reduction in transport costs is expressed in percentage points. In the world including additive costs, we simulate the observed reduction in $\tau$ and iterate on the reduced value of $t$ such that the simulated value of $t/\widetilde{p}_x^{\text{fas}}$ matches the empirical target.  }}}
\end{table}

In line with the literature, we find that trade liberalization induces welfare gains whatever the transport mode (and the associated magnitude of the transport costs decrease). The originality of our results rather dwells on comparing the welfare gains depending on the type of cost reduction. On that side, the welfare gains are higher when the trade cost reduction occurs through a decrease in the additive component. This conclusion stands in accordance with the analytical results of \cite{sorensen2014}. This is consistent with the analytical results of \cite{sorensen2014}, in a different exercise of trade liberalization though. \footnote{\cite{sorensen2014} compares the welfare effect of a trade costs reduction depending whether it is exclusively ad-valorem or exclusively additive, the comparison between both trade cost decreases being a criterion of equal impact on trade openness. Our quantitative model offers a more general and flexible set-up, as we can virtually consider all cases of trade liberalization. Based on our empirical estimates, we compare the effects of the same decrease in overall transport costs depending on 1) whether additive costs are present or not, and 2) when they are present, assuming that the trade cost decrease is achieved by a reduction of both additive and ad-valorem components.} Quantitatively, the welfare gains are sizeable. The reduction variable transport costs generates a relative welfare gain higher by around 1.1 percentage point when such a reduction occurs through the additive component (which accounts for 50 to 60\% of the total cost decrease, depending on the transport mode), rather than when it realizes through a decrease in iceberg costs only (Table \ref{tab:resultats_modele}, panel (1), Column (B) vs (A) for Air, (D) vs (C) for Vessel). \cite{Irrazabal_2015} also quantify the welfare gains of a trade cost decrease in presence of additive costs. In contrast to \cite{Irrazabal_2015} though, our empirical analysis allows us to simulate the effects of the reduction in transport costs actually observed in the data over the period 1974-2019, in all their dimensions (additive and ad-valorem). \smallskip

@@

To understand the changes induced by the additive costs, we start considering the benchmark model \textit{\`{a} la M\'{e}litz} with only ad-valorem costs (Table \ref{tab:resultats_modele}, Columns (A) and (C)). As can be inferred from Table \ref{tab:resultats_modele}, panel (1), the reduction in international trade costs is driven by positive competitive and variety effects (whatever the transport mode). The positive variety effect can be understood as the result of two opposite forces. On the one hand, if we consider the reduction in the iceberg cost only (constant $f_x$, as reported in Panel (2)), the variety effect is negative. With the reduction in $\tau$, more firms can now export, but this also increases the expected average profit. This raises the productivity threshold to enter the domestic market. Accordingly, the number of domestic active firms reduces, even if a larger number of them is able to export. As a whole, the total number of varieties accessible to the consumer reduces, which reduces welfare everything else equal. The reduction in the number of available varieties for the household stands in line with the insights of the canonical \cite{melitz} model. On the other hand, the reduction in the fixed export cost pushes more firms to export everything else equal. This increases the number of available imported varieties for the consumer. This more than counteracts the reduction in the number of domestic varieties, inducing a positive variety effect when considering both the reduction in $\tau$ and in $f_x$.

The positive price competitiveness effect can be accounted for as follows. Due to the tougher selection on the domestic market, the less productive firms no longer enter. As a result, the average productivity on the domestic market increases, which pushes the average productivity upwards. As reported in Table \ref{tab:resultats_modele}, Panel (1), under the joined reduction in $\tau$ and $f_x$, both the price competitiveness and the variety effects induce positive welfare gains, consistently with the analytical results of \cite{melitz}.

What does it change when the reduction in international transport costs also occurs through a reduction in its additive component? As can be inferred from Table \ref{tab:resultats_modele}, the stronger welfare gains of trade liberalization are driven by a more positive competitiveness effect, which more than offsets a variety effect which is now welfare-detrimental, whether $f_x$ remains constant or reduces.

Consider first the price competitiveness effect. As underlined by \cite{sorensen2014}, a reduction in the additive cost reduces the export-market price of the high-productivity exporters relative to the low-productive ones (as can be inferred from Equation (\ref{eq:px})). Within the exporters category, this favors the more productive firms, whose market share, e.g. revenues, increases relatively more (see Equation (\ref{eq:rx})). Everything else equal, this pushes the average productivity on the export market upwards. Unlike pure ad-valorem costs, the reduction in additive costs induces a \textit{within-exporters selection effect}, in favor of the more productive firms. At the general equilibrium, this \textit{within-exporters selection effect} also affects the \textit{classic} selection effect at the extensive margin at the heart of the \cite{melitz} model. As the expected export profit increases, it is necessary to be more productive to satisfy the free-entry condition. Everything else equal, the threshold productivity on the domestic market rises, as well as the average productivity. Accordingly, the competitiveness effect increases by more when the trade liberalization involves a reduction in the additive component.

In contrast with the only ad-valorem cost case, the variety effect turns out to be negative, whether $f_x$ is kept constant or reduced. This can be accounted for as follows. As discussed above, the reduction in additive costs favors the market share of the more productive firms within the set of exporters. This makes entry on the export market tougher, thereby moderating the increase in the number of exported varieties in comparison with the pure ad-valorem case. This is yet counteracted by the reduction in the export fixed cost, such that the number of exported varieties ultimately raises. The negative variety effect is hence mostly attributable to a reduction in the number of domestic incumbents. As the selection on the domestic market is more stringent, the mass of total active firms indeed reduces by a larger amount. From the consumer side, this induces a negative variety effect that is welfare-detrimental everything else equal. The strength of the (larger) competitiveness effect still dominates, inducing positive welfare gains, and even larger when the trade liberalization involves a reduction in additive costs.\medskip

As mentioned above, our results dwell on the joined reduction in both the variable and the fixed export costs, based on the empirical evidence provided in Section  \ref{sec:basic_results} and in \cite{Lincoln_McCallum2018}. One might yet wonder about the welfare gains induced by the reduction in variable costs only, i.e. \textit{ceteris paribus} for a given value of the fixed export cost. The answer can be found in Table \ref{tab:resultats_modele}, Panel (2). In this case, the fixed export cost is kept constant at its initial value, the only structural change being the decrease in variable transport costs (either ad-valorem only or both additive and ad-valorem). Two results emerge. First, the welfare gains induced by the trade liberalization are lower in magnitude. While this is not a surprising result \textit{per se} (since we only consider the reduction in one source of international frictions, rather than two), this exercise yet delivers a quantification of the welfare gains associated with a reduction in the fixed export cost: They are substantial, by an amount between 0.6 to 1.25 percentage point depending on the transport mode. Second, the main message of the paper remains valid. The welfare gains of the trade liberalization are still higher when part of this structural change is achieved through a reduction in additive transport costs, by the same order of magnitude (+ 1.2 percentage point). Put differently, the reduction in the fixed export cost does not seem to depend on the composition of the variable costs. In the end, the welfare gains are the highest when trade liberalization induces a combined reduction in both variable and fixed export costs, and when (part of) the reduction in the variable part is achieved through a decrease in the additive component, as observed empirically.\smallskip

In line with \cite{sorensen2014} and \cite{Irrazabal_2015}, our overall theoretical results point out the importance of taking into the additive components of international trade costs. Additive costs have a substantial impact on the extensive margin of trade, by affecting the composition of the basket of exported goods. This, in turn, has important consequences in normative terms. Relying on our empirical estimates of transport costs over 1974-2019, we obtain that the 3.4\% reduction of variable costs in Air transport, combined with the reduction in the fixed export costs, has induced an 3.16\% increase in welfare, much higher than if only iceberg costs were modeled (2.5\%). A similar result applies with the observed 6.4\% in total variable costs in maritime transport along with the reduction in the fixed export costs, with a relative increase in welfare by 4.3\% (as compared to 3.6\% in a pure ad-valorem world). Neglecting additive costs thus leads to a substantial underestimation of the welfare gains of trade liberalisation. While we have obtained these results in the canonical \cite{melitz} model, they call for integrating additive costs in New Trade models in view of better assessing the (disputed) question of the welfare gains of trade liberalization.

@@

From this, we deduce the recursive link: $\widehat{t}_{ist} = \widehat{t}_{is74} \times \exp(\gamma_t)$.

Given the constraint $t>0$, we then obtain the percentage change from 1974 from $\Gamma^{add}_{ist} = 100\frac{\widehat{t}_{ist}}{\widehat{t}_{ik74}} = 100\exp(\gamma_t)$. Again, it is independent of the product-origin country pair. We can thus rewrite the time-trend series for the additive transport cost component as:
\begin{equation}
\Gamma^{add}_t  = 100\exp(\gamma^{add}_t) \label{eq:tcadd_compoeffect}
\end{equation}
